---
title: "Confidence Interval for Population Mean"
author: "Dieter Haage"
date: "August 1, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Step 1) Load Data
```{r echo=FALSE}
lifeexp <- read.csv('Life_exp_non_OECD_subset.csv', header = T)

View(lifeexp)

str(lifeexp)
```

# Step 2) Perform Exploratory Analysis

```{r}
qqnorm(lifeexp$Male)

```
Data appears normal, given small sample size

# Step 3) Remove outliers

```{r}
boxplot(lifeexp$Male)
```

No extreme outliers, hence no points are removed 

# Step 4) Compute sample mean, SD and SE of the mean

```{r}
n <- nrow(lifeexp)

age.mean <- mean(lifeexp$Male)
age.std <- sd(lifeexp$Male)
age.SE <- age.std/sqrt(n)

print(paste("The mean is", signif(age.mean,4),'with standard deviation of', signif(age.std,4),"and standard error of",signif(age.SE,4)))

```

# Step 5) Compute the t-value of the 95% confidence interval
```{r}

degrees.freedom <- n-1
alpha <- 0.05
tstar <- qt(1- alpha/2, df = degrees.freedom)

print(paste("tstar is", signif(tstar,4)))


```

# Compute Confidenece Interval
```{r}
age.min <- age.mean - tstar*age.SE
age.max <- age.mean + tstar*age.SE

print(paste("The confidence interval is ", signif(age.min,4),"< age <", signif(age.max)))


```

# Step 7) Interpret Findings

Based on the sample results, we are 95% confident that the limits of 60.72 and 69.62 contain the population mean life expectancy of males from non-OECD countries. Interestingly, the population mean male life expectancy of OECD countries' value of 77.86 lies beyond this confidence interval, suggesting that male life expectancy is less in non-OECD countries than in OECD countries..

 
